\textbf{\supplement}\vskip 2ex \mytitle\vskip 2ex

### 2022-02-28 paper-seb-supplementary-material.tex
All files updated for PLOS Feb 28 2022 submission

\PassOptionsToPackage{hyphens}{url}
\documentclass{article}
\renewcommand\thesubsection{\thesection.\alph{subsection}}
%\usepackage{hyperref}
%\usepackage{refcount}
\usepackage{url}
\usepackage{doi}
\usepackage{longtable}
\usepackage{color}
\usepackage{amsmath,amssymb}
\def\httpURL#1{\href{http://#1}{\textcolor{blue}{#1}}}
\def\LONGhttpURL#1#2{\href{http://#1}{\textcolor{blue}{#2}}}

\input paper-seb-macros.tex

\makeatletter
      \renewcommand{\thefootnote}{\arabic{footnote}}
      {\let\oldbibcite=\bibcite
      % \bibcite{bad-code}{{31}{2020}{{Richards and Boudnik}}{{Richards and Boudnik}}}
      \newcount\mainNumberOfReferences \mainNumberOfReferences=0
      \def\bibcite#1#2{\global\advance\mainNumberOfReferences by 1\oldbibcite{#1}{#2}}
      \input paper-seb-main.aux
      %\inputifexists{paper-seb-main.aux}{You need to run Latex on paper-seb-main.tex because we need its aux file}
      }
\makeatother

@@

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

@@

\maketitle

@@

\makeatletter
\def\globalLabel#1{% we also need journal code-policies to be used in the main paper...
\immediate\write\infofile{\bslash newlabel{#1}{{\arabic{section}}{\thepage}{DUMMYA}{DUMMYB}{DUMMYC}}}
% as well as in the current document....
\immediate\write\@auxout{\string\newlabel{#1}{{\arabic{section}\dotalph{subsection}}{\thepage}{AAA}{BBB}{CCC}}}
}
\makeatother

% start supplementary material's page counters to continue on from the paper's page numbering
%\setcounterpageref{page}{LastPage}
%\addtocounter{page}{1} % otherwise we start with the same page that the main paper ended on (we want to be on the next page:-)!

@@

%\renewcommand{\thesubsection}{\alph{subsection}}
\def\sectioncontents#1{\immediate\write\contentsfile{\bslash contents{\arabic{section} #1}{\thepage}}}
\def\subsectioncontents#1{\immediate\write\contentsfile{\bslash subcontents{\alph{subsection}\ #1}{\thepage}}}
\def\starredsubsectioncontents#1{\immediate\write\contentsfile{\bslash subcontents{#1}{\thepage}}}

\noindent\vbox{% The table of contents
\def\contents#1#2{\vskip 0mm \noindent{\bfseries #1} \dotfill\ #2\\}
\def\subcontents#1#2{\hbox to 1em{}{#1} \dotfill\ #2\\}
\noindent\input generated-supplementary.toc
%\inputifexists{generated-supplementary.toc}{you need to Latex supplementary-material.tex again}
}
\newwrite\contentsfile
\immediate\openout\contentsfile=generated-supplementary.toc

@@

\vbox{\noindent\input generated-minicontents.tex
%\inputifexists{generated-minicontents.tex}{You need to run Latex again on supplementary-material.tex to generate the minicontents file}
}

@@

@@

There is no fundamental difference between code and data, and no distinction that is relevant for scientific publication purposes. There is no distinction one can imagine that cannot easily, even accidentally, be circumvented. In other words, a journal's data policies and code policies should be the identical --- and the conventionally stricter data policies should also apply to code. It is baffling that some journals have data policies that are weaker than their data policies; it is certainly indefensible to have no code policies at all.

@@

This section's discussion may sound like arcane and irrelevant pedantry, but these issues are at the very foundations of Computer Science.\footnote{{Many of the foundational issues were explored thoroughly by Christopher Strachey and others in the 1960s; Strachey's classic lectures are reprinted in an accessible 2000 publication \cite{strachey}. Being originally a very old paper this classic introduction is much easier to read than many more recent discussions of the foundations of Computer Science.}} If we ignore or misunderstand these basic things --- or overlook them in policies and procedures --- bugs are the inevitable (and confusing) consequence.

The main paper points out that data is often embedded in code as ``magic numbers.'' Let's now explain how.

A fragment of program code might say

\begin{center}\texttt{x = 324.9+sin(theta*pi/2);}\end{center}

This is clearly all source code, but the number 324.9 above is likely to be some sort of relevant data, though it might be a physical constant whose value does not depend at all on \emph{this\/} experiment. The next hard-coded value mentioned in the calculation is difficult to categorize: is the value of $\pi$ empirical data or is it part of a standard formula? Historically, even $\pi$ was definitely an empirical value, but today it is a mathematical constant --- except in experiments to determine its value empirically. The point is, the distinctions between data, program and even mathematical constants are purely a matter of perspective.

Unfortunately, there is data that is extremely easy to overlook (and therefore is very hard to manage). You may assume that the function \texttt{sin}, as used in the calculation example above, is the standard trigonometric function for calculating sines (and because of $\pi$, you assume it is taking radians as the parameter type) but almost all programming languages allow \texttt{sin} to be any function whatsoever. Confusingly, it is generally a different function when the code is run on a different computer.

It is impossible to tell. More complex code will often have many facts ``hard wired'' into the code --- so in fact the code contains data. Code can even read in formulas from data and compile them to perform further calculations, and so on. Equally, data can control the flow of code. For example, data summarizing patients may include their gender, but the program processes males and females differently. Then data starts becoming like code.

{Many computer programs blur the distinctions deliberately, to create virtual machines. Data is then run on the virtual machine as program. Many programs provide standard features to do this, such as LISP's and JavaScript's \texttt{eval} functions. Henderson's book \cite{henderson} builds an elegant Pascal program to run \emph{any\/} LISP program as data, and then shows that the LISP program can run itself running other programs, so it is now its own code and \emph{its\/} data --- despite being purely data to the Pascal program. There are numerous advantages to doing this, including: the Pascal program is not just reading data, but structured data that must conform to the rules of LISP; the LISP running itself runs faster than the original Pascal running LISP, even though the Pascal virtual machine is still doing it in the recursive case; LISP is a much more powerful language than Pascal, so a virtual machine can be used to escape the barriers of a limited implementation such as Pascal. In short, any distinctions between code and data are impossible to maintain.}

In the present paper, we knowingly built on this blur between data and code. However, what we did was not unusual except in our explicit and rigorous approach to managing and summarizing data reliably in the paper.

@@

@@

There is considerable debate over what good commenting practice is, but this is because comments have many roles --- from helping students to get marks in assessments, asserting intellectual rights, reminding the developer of things to do, managing version control, to explaining the code to third parties. Different programming languages also develop ``cultures'' that encourage different approaches for comments (examples include R Markdown, Mathematica Notebooks, JavaDoc, Haskell's Haddock, and so on). For scientific code, however, the explanatory role is critical, and this is what was assessed in this survey. Note that comments used for explanation does not preclude any other uses.

The completeness or executability of code was not assessed, although if code was obviously incomplete this was noted. Whether code runs as claimed is a matter of research integrity, which is beyond the scope of this survey. What is relevant to the study is whether the code is described in sufficient detail that the methods used can be scrutinized. Obviously being able to run the code will help, but clarity in documentation and comments is critical. It is more like ``can we see the critical pages from your lab book so we understand what you did?'' rather than ``can we have free run of your laboratory?''

@@

\input generated-example-data.tex

The data was written out by hand, after reading and reviewing each paper in the survey. In total there are \countFields\ data fields available for documenting papers, but not all need be used for each paper; for example, the field \texttt{hasCodeTested} defaults to \texttt{false}, so it need not be set --- it is also an error to set it if another field asserts there is no code to evaluate! (A separate JSON data structure maps the data fields to English descriptions, along with default values if they are optional descriptors.)

@@

where {\tt\small \bslash plural} automatically writes a word (``paper'' in this case) in singular or plural form as required (in this case, the variable \texttt{countUsesVersionControlRepository} = \the\countUsesVersionControlRepository). When typeset the text above is generated with the relevant number or numbers inserted from the JSON data.

Some of the files generated from the JSON data are Unix shell scripts. For example, details of all the papers with GitHub repositories are automatically collected into a shell script so the repositories can be cloned locally and then measured to generate table \ref{table-repo-summary}, as used in the main paper.

@@

\highlightmysubsection{Detecting and defending against error}
\label{detecting-against-error}
It must be emphasized that an automatically-guaranteed same number appearing in multiple different contexts is an extremely effective way of defending against common Human Factors errors.

% check it works for other numbers 
%\countUsesVersionControlRepository=296

Normally, when we write \the\countUsesVersionControlRepository\ in a paper, we proof read it as ``just a number'' and the sentence would likely seem to make as much sense if the number was 
\newcount\t
\t=\countUsesVersionControlRepository
\divide \t by 10
\newcount\fraction
\fraction=\countUsesVersionControlRepository
\newcount\tent \tent=\t \multiply \tent by 10
\advance\fraction by -\tent
\the\t.\the\fraction,
\t=\countUsesVersionControlRepository
\advance \t by -1
\the\t,
\t=\countUsesVersionControlRepository
\advance \t by 1
\the\t, or
\t=\countUsesVersionControlRepository
\multiply \t by 10
\the\t\ --- we hardly bother to pay attention because we know what we are reading; at least we know what we meant to write. It is very hard to spot one's own typos.

@@

All the discussion and examples above were generated automatically, and have been checked  correct for other base values than \the\countUsesVersionControlRepository. This approach, too, considerably helps defend against common Human Factors errors. For example, if we set \texttt{\bslash countUsesVersionControlRepository}=\the\countUsesVersionControlRepository\ to be 2348, say, then all of the sentences say something unexpected and so have to be more carefully proof-read, significantly reducing confirmation bias.

If any of the numbers used in a paper were safety critical (e.g., lives directly depend on their correct values) then further checks would have been made to help detect and avoid errors. For example, \LaTeX\ makes it very easy to check that numbers fall within reasonable ranges, or have any other required safety properties. For the present paper, a potential problem is that the paper is typeset \emph{before\/} the JSON data has been analyzed; in which case, none of the variables, like \texttt{\bslash countUsesVersionControlRepository}, will have been correctly set and their values could be undefined or nonsense.

\newcount\min \min=5
\newcount\max \max=20
%\countUsesVersionControlRepository=4
\ifnum \countUsesVersionControlRepository>\max countUsesVersionControlRepository is far too large! \fi
\ifnum \countUsesVersionControlRepository<\min countUsesVersionControlRepository is far too small! \fi
\ifnum \countUsesVersionControlRepository>\max
	\else 
		\ifnum \countUsesVersionControlRepository<\min
			\else
								\textbf{This is \emph{automatic\/} confirmation that $$\the\min \leq \mbox{\tt countUsesVersionControlRepository} \leq \the\max$$ and therefore falls within pre-defined sanity limits for this paper.}
		\fi
\fi

@@

@@

It is noteworthy that none of the journals sampled permit any reliable style of managing data in published papers, such as described above in section \ref{detecting-against-error}. In particular, for all the papers that had accessible code, the code included explicit (and relevant) data that was not archived \emph{as\/} data in the journal repositories.
\\

{\sf
\noindent\textbf{Extract from \emph{Royal Society Open Science\/} author guidelines} \\
--- It is a condition of publication that authors make the primary data, materials (such as statistical tools, protocols, software) and code publicly available. These must be provided at the point of submission for our Editors and reviewers for peer-review, and then made publicly available at acceptance. [\ldots] As a minimum, sufficient information and data are required to allow others to replicate all study findings reported in the article. Data and code should be deposited in a form that will allow maximum reuse. As part of our open data policy, we ask that data and code are hosted in a public, recognized repository, with an open licence (CC0 or CC-BY) clearly visible on the landing page of your dataset.

\vskip 3mm
{\raggedleft\scriptsize
\url{royalsociety.org/journals/authors/author-guidelines/#data} \\
accessed 29 July 2020.\par
}

\noindent\textbf{Extract from \emph{Nature Digital Medicine\/} author guidelines}\\
--- A condition of publication in a Nature Research journal is that authors are required to make materials, data, code, and associated protocols promptly available to readers without undue qualifications. [\ldots] A condition of publication in a Nature Research journal is that authors are required to make unique materials promptly available to others without undue qualifications.

{\vskip 3mm\scriptsize
\raggedleft
\url{www.nature.com/nature-research/editorial-policies/reporting-standards#availability-of-data} \\
accessed 29 July 2020.\par
}

\noindent\textbf{Extract from \emph{Lancet Digital Health\/} author guidelines}\\
--- Has detailed data policies, but no code policy.
{\vskip 3mm
\raggedleft\scriptsize
\url{marlin-prod.literatumonline.com/pb-assets/Lancet/authors/tldh-info-for-authors.pdf}
\\
accessed 29 July 2020.\par
}

\noindent\textbf{Extract from \emph{Journal of Vascular Surgery\/} author guidelines}\\
--- Has detailed data policies, but no code policy. While no \emph{Journal of Vascular Surgery\/} papers were surveyed, the following statement on data policies is relevant for section \ref{jvs-policy} in the main paper:

>>>>

\iftrue
    \PassOptionsToPackage{hyphens}{url}
    \documentclass[10pt,letterpaper]{article}

    \usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
    % Use adjustwidth environment to exceed column width (see example table in text)
    \usepackage{changepage}

    % line numbers
    \usepackage[right]{lineno}
    \linenumbers

    \renewcommand\thesubsection{\thesection.\alph{subsection}}
    %\usepackage{hyperref}
    %\usepackage{refcount}
    \usepackage{url}
    \usepackage{doi}
    \usepackage{longtable}
    \usepackage{color}
    \usepackage{amsmath,amssymb}
    \usepackage{fancyhdr}
    \def\httpURL#1{\href{http://#1}{\textcolor{blue}{#1}}}
    \def\LONGhttpURL#1#2{\href{http://#1}{\textcolor{blue}{#2}}}

    \input paper-seb-macros.tex
\fi

\makeatletter
      \renewcommand{\thefootnote}{\arabic{footnote}}
      {\let\oldbibcite=\bibcite
      % example: \bibcite{bad-code}{{31}{2020}{{Richards and Boudnik}}{{Richards and Boudnik}}}
      \newcount\mainNumberOfReferences \mainNumberOfReferences=0
      \def\bibcite#1#2{\global\advance\mainNumberOfReferences by 1\oldbibcite{#1}{#2}}
      \input paper-seb-main.aux 
      \resetlinenumber[\MainPaperLastLine]
      }

\makeatother

@@

@@

\maketitle
\pagestyle{fancy}

% pageref gives us a hyperlink which looks unsightly
\makeatletter
\def\rememberpage{\immediate\write\@auxout{\bslash gdef\bslash SupplementalLastPage{\thepage}}}
\makeatother

\rfoot{\supplement\ \thepage/\SupplementalLastPage}
\lfoot{\today}
\cfoot{}
\lhead{}
\rhead{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\thispagestyle{empty}

@@

\makeatletter
\def\globalLabel#1{% we also need journal code-policies to be used in the main paper...
\immediate\write\infofile{\bslash newlabel{#1}{{\arabic{section}}{\thepage}{DUMMYA}{DUMMYB}{DUMMYC}}}
% as well as in the current document....
\immediate\write\@auxout{\bslash newlabel{#1}{{\bslash arabic{section}\dotalph{subsection}}{\thepage}{AAA}{BBB}{CCC}}}
}
\makeatother

% start supplementary material's page counters to continue on from the paper's page numbering
\setcounterpageref{page}{LastPage}
\addtocounter{page}{1} % otherwise we start with the same page that the main paper ended on (we want to be on the next page:-)!

@@

\def\sectioncontents#1{\immediate\write\contentsfile{\bslash contents{\arabic{section} #1}{\thepage}}}
\def\subsectioncontents#1{\immediate\write\contentsfile{\bslash subcontents{\alph{subsection}\ #1}{\thepage}}}
\def\starredsubsectioncontents#1{\immediate\write\contentsfile{\bslash subcontents{#1}{\thepage}}}

\noindent\vbox{% The table of contents
\def\contents#1#2{\vskip 0mm \noindent{\bfseries #1} \dotfill\ #2\\}
\def\subcontents#1#2{\hbox to 1em{}{#1} \dotfill\ #2\\}
\noindent
\input generated-supplementary.toc
}

\newwrite\contentsfile
\immediate\openout\contentsfile=generated-supplementary.toc

@@

\vbox{\noindent
      \input generated-minicontents.tex
      }

@@

\begin{change}Interestingly, the only paper reviewed that claimed to do any independent testing \cite{onlyPaperWithChecks} failed to include any testing in its data or code repository, so the testing itself --- the essential quality assurance of the code --- is not open to scrutiny (e.g., the code and the ``independent'' code are likely to contain common code, data, and common bugs).\end{change}

@@

\begin{change}
All computer systems are in principle equivalent to Turing Machines, and Turing Machines make no distinction between program and data. It is possible to define Turing Machines that do separate program code and data, but as soon as a Universal Turing Machine is constructed, its data \emph{is\/} code. Indeed, Universal Turing Machines are a theoretical abstraction of virtual machines, which are used widely in practical computing. Java, for instance, runs in a virtual machine, so any Java program code (and any data it uses) is in fact merely \emph{all\/} data to the Java virtual machine. At another extreme, $\lambda$-calculus is purely program source code, yet $\lambda$-calculus is equivalent to Turing Machine computation. Therefore, even the ``pure'' programs of $\lambda$-calculus also represent data.
\end{change}

These elementary theoretical considerations underly an important practical fact: there is no fundamental difference between code and data, and no distinction that is relevant for scientific publication purposes.

There is no code/data distinction one can imagine that cannot easily, even accidentally, be circumvented. In other words, a journal's data policies and code policies should be the identical --- and the conventionally stricter data policies should also apply to code. It is baffling that some journals have data policies that are weaker than their data policies; it is certainly indefensible to have no code policies at all.

@@

This section's discussion may sound like arcane and irrelevant pedantry, but these issues are at the very foundations of Computer Science.\footnote{{Many of the foundational issues were explored thoroughly by Christopher Strachey and others in the 1960s; Strachey's classic lectures are reprinted in an accessible 2000 publication \cite{strachey}. Being originally a very old paper this classic introduction is much easier to read than many more recent discussions of the foundations of Computer Science.}} If we ignore or misunderstand these basic things --- or overlook them in policies and procedures --- bugs and irreproducibility are the inevitable (and confusing) consequence.

The main paper points out that data is often embedded in code using ``magic numbers.'' Let's now explain how.

A simple fragment of program code might say

\newcount \magicNumber \magicNumber = 324
\begin{center}\texttt{x = \the\magicNumber+sin(theta*pi/180);}\end{center}

This is clearly all source code, but the number \the\magicNumber\ above is likely to be some sort of relevant data, though it might be a physical constant whose value does not depend at all on \emph{this\/} experiment. The next hard-coded value mentioned in the calculation is difficult to categorize: is the value of $\pi$ empirical data or is it part of a standard formula? Some programming languages like Mathematica treat $\pi$ as an exact mathematical constant (e.g., Mathematica calculates $\tan\pi/4 = 1$ exactly), but $\pi$ is \emph{also\/} definitely an inexact empirical value.\footnote{A record set on 19 August 2021, the most accurate value of $\pi$ then known was 62,831,853,071,796 digits %as calculated by the University of Applied Sciences, Chur, Switzerland 
\url{www.fhgr.ch/en/specialist-areas/applied-future-technologies/davis-centre/pi-challenge}}

The point is, the distinctions between data, program and even mathematical constants are purely a matter of perspective.

Unfortunately, there is data that is extremely easy to overlook (and therefore very hard to manage) because it is embedded in arbitrary ways in code. You may assume that the function \texttt{sin}, as used in the calculation example above, is the standard trigonometric function for calculating sines (and because of the $\pi$ in the expression, you assume \texttt{theta} is degrees and \texttt{sin} is taking radians as its parameter type) but almost all programming languages allow \texttt{sin} to be any function whatsoever. Confusingly, even if it calculates sines, it is generally a different function when the code is run on a different computer producing numbers that are not exactly the same.

It is impossible to tell.

\begin{change}
\subsection{When magic numbers become magic code}
Data often controls the flow of code. For example, data summarizing patients may include their gender, but the program processes males and females differently. Then data becomes code.

Arbitrary numbers appearing in code are obviously magic numbers, but code often conceals the magic numbers of data by ``programming them away'' during the coding process.

For example, the magic number \the\magicNumber\ was explicit in the line of code shown above, but if somewhere else the program says

\begin{center}\texttt{if evenQ(\the\magicNumber) then A; else B;}\end{center}

many programmers would optimize this to \texttt{A}, because they know the condition is true because of their assumptions. This now seems to be a more efficient program because it has avoided a test (which a modern complier would have optimized away anyway). Unfortunately, the previously explicit dependency of the code on the magic number \the\magicNumber\ has completely disappeared.

Obviously this example seems trivial, but it illustrates that programmers do some of their programming while writing code, and many assumptions disappear completely and have no representation in the final code. More complex code will have many facts hard wired into the code --- so in fact the code contains data. Code can even read in formulas from data and compile them to perform further calculations, and so on.

This is one reason bugs --- effectively incorrect assumptions --- are so hard to find, because they have no concrete form in the final program.

\subsection{When data is code}
{Many computer programs blur the simplistic code/data distinctions deliberately, to create virtual machines. Data is then run on the virtual machine as program. Many programs provide standard features to do this, such as LISP's and JavaScript's \texttt{eval} functions. Henderson's book \cite{henderson} builds an elegant Pascal program to run \emph{any\/} LISP program as data, and then shows that the LISP program can run itself running other programs, so it is now its own code and \emph{its\/} data --- despite being purely data to the Pascal program. There are numerous advantages to doing this, including: the Pascal program is not just reading data, but structured data that must conform to the rules of LISP; the LISP running itself runs faster than the original Pascal running LISP, even though the Pascal virtual machine is still doing it in the recursive case; LISP is a much more powerful language than Pascal, so a virtual machine can be used to escape the barriers of a limited implementation such as Pascal. In short, any distinctions between code and data are impossible to maintain.}

AI and Machine Learning are further examples of exploiting data as code. Typically a program learns from a training set of data, and then processes future data differently depending on what it has learned. In other words, the original data becomes a model which is now code.

\subsection{Exploiting code as data for more reliable science}
\end{change}\noindent
In the present paper, we knowingly built on this blur between data and code. However, what we did was not unusual except in our explicit and rigorous approach to managing and summarizing data reliably in the paper.

@@

\begin{change}\sequestion{What evidence is there of rigorous (e.g., code review and tool-based) independent methods being used to manage coding bias?}\end{change}

@@

\begin{change}
Although the pilot survey is not a systematic review, following Preferred Reporting Items for Systematic Reviews and Meta-Analyses (PRISMA) \cite{prisma} it is good practice to disclose details of the reviewers. In the present case, the selected papers were assessed by the author of the present paper. The study was not blinded. This is, of course, a limitation of the study. However, the reviewer is a full professor of computer science, who has taught and assessed computer software since the 1970s, using moderated and peer-reviewed processes for undergraduate and postgraduate computing degrees, and is well aware assessing code quality has been a lively topic in software engineering for decades (there is now international standard ISO/IEC 9126, updated to ISO 25000). The author has written legal documents analyzing software for criminal cases involving faulty software. The author has approximately 
\newcount \papers \papers=340 % peer reviewed
\advance \papers by 188 %invited
\the\papers\ published papers in computer science.

The evaluations performed for the present paper are at a trivial level where sources of bias should have a negligible effect, particularly given that the overall conclusion is consistent across both the diverse sample and the computational science-based papers cited in the paper that did not form part of the selected sample. As said in the body the paper: \emph{the fact that the specifically identified problems are elementary errors in software engineering (see the discussion in \emph{[main paper]\/} section \ref{section-discussion}) suggests more sophisticated analysis is not required}.

In any case, as stated in the main paper, the full dataset and analysis code is available at \begin{center}\url{github.com/haroldthimbleby/Software-Enginering-Boards}\end{center} and the reviewed papers are (unless retracted) still available online for independent assessment. %The data is also provided in full later in this \supplement.
\end{change}

There is considerable debate over what good commenting practice is, but this is because comments have many roles --- from helping students to get marks in assessments, asserting intellectual rights, reminding the developer of things to do, managing version control, to explaining the code to third parties. Different programming languages also develop ``cultures'' and tool-based systems that encourage different approaches for comments (examples include R Markdown, Mathematica Notebooks, JavaDoc, Haskell's Haddock, and so on). For scientific code, however, the explanatory role is critical, and this is what was assessed in the present survey. It is notable that no such tool-based approach to code or documentation was used in any code reviewed.

The completeness or executability of code was not assessed, although if code was obviously incomplete this was noted. Whether code runs as claimed is a matter of research integrity, which is beyond the scope of this survey. What is relevant to the study is whether the code is described in sufficient detail that the methods used can be scrutinized. Obviously being able to run the code will help, but clarity in documentation and comments is critical. It is more like ``can we see the critical pages from your lab book so we understand what you did?''\ rather than ``can we have a free run of your laboratory, even though we don't understand the details of the science?''

@@

\newdimen \codeIndent
\codeIndent = 1em
\input generated-example-data.tex

The data was entered by hand (as JSON terms), after reading and reviewing each paper in the survey. In total there are \countFields\ data fields available for documenting papers, but not all need be used for each paper; for example, the field \texttt{hasCodeTested} defaults to \texttt{false}, so it need not be set --- it is also an error to set it if another field asserts there is no code to evaluate! (A separate JSON data structure maps the data fields to English descriptions, along with default values if they are optional descriptors.)

@@

where {\tt\small \bslash plural} automatically writes a word (``paper'' in this case) in singular or plural form as required.

\begin{change}For each of the \plural{\dataVariableCount}{variable} used in the paper, JavaScript generates a \LaTeX\ header file that declares and assigns a calculated value. The header file is then included in the paper using \LaTeX's standard \texttt{\textbackslash input} command. Here is an example of one such automatic definition:

\begin{verse}\tt
\textbackslash newcount \textbackslash dataVariableCount \\
\textbackslash dataVariableCount = \the\dataVariableCount
\end{verse}

so the named value is then available for the author to use when the paper is typeset. The tables generated for the present paper also include data that did not require names for explicit use elsewhere in the \LaTeX\ files, so these numbers are not counted in the specific \the\dataVariableCount\ variable count.
\end{change}

Some of the files generated from the JSON data are Unix shell scripts. For example, details of all the papers with GitHub repositories are automatically collected into a shell script so the repositories can be cloned locally and then measured (using \texttt{awk} scripts), e.g., to generate table \ref{table-repo-summary}, as used in the main paper.

@@

\highlightmysubsection{Detecting and defending against error}
\label{detecting-against-error}
% check it works for other numbers 
%\countUsesVersionControlRepository=296

Normally, when we write a number like \the\countUsesVersionControlRepository\ in a paper, especially longer or more complex numbers, we will later proof read them as ``the numbers we intended to write'' --- as remembering what we meant is easier than reading the details. Unfortunately, a sentence would likely seem to make as much sense when a number has been erroneously typed as, say,
\newcount\t
\t=\countUsesVersionControlRepository
\divide \t by 10
\newcount\fraction
\fraction=\countUsesVersionControlRepository
\newcount\tent \tent=\t \multiply \tent by 10
\advance\fraction by -\tent
\the\t.\the\fraction,
\t=\countUsesVersionControlRepository
\advance \t by -1
\the\t,
\t=\countUsesVersionControlRepository
\advance \t by 1
\the\t, or
\t=\countUsesVersionControlRepository
\multiply \t by 10
\the\t\ --- we hardly bother to pay attention because we \begin{change}think we\end{change} know what we are reading; at least we know what we meant to write. Worse, the more often we proof read a document, the more we remember, so the better we know what we think we said, and the more casual our proof reading becomes. It is very hard to spot all of our own typos.

@@

All the discussion and examples above were generated automatically, and have been checked correct for other correct values than \the\countUsesVersionControlRepository. This approach, too, considerably helps defend against common Human Factors errors. For example, if we set \texttt{\bslash countUsesVersionControlRepository}=\the\countUsesVersionControlRepository\ to be 2348, say, then all of the subsequent sentences that mention it will say something unexpected and so have to be more carefully proof-read, significantly reducing confirmation bias. \begin{change}The approach turns a possibly-hard-to-spot \emph{single\/} error into \emph{multiple\/} errors spread throughout the paper into different contexts, thus increasing the chances of noticing the error.

It must be emphasized that an automatically-guaranteed number that is supposed to be the same appearing in multiple different contexts is an extremely effective way of defending against common Human Factors errors. As the number is proof read, the different contexts encourage it to be read more carefully, and in different ways.
\end{change}

If any of the numbers used in a paper were safety critical (e.g., lives directly depend on their values) then further checks would have been made to help detect and avoid errors. \LaTeX\ itself makes it very easy to check that numbers fall within reasonable ranges, or to have any other required safety properties. For the present paper, a potential problem is \begin{change}if\end{change} the paper is mistakenly typeset \emph{before\/} the latest JSON data has been analyzed; in which case, none of the variables, like \texttt{\bslash countUsesVersionControlRepository}, will have been correctly set and their values could be undefined or nonsense \begin{change}(e.g., from a debugging run of \texttt{data.js}). Variables might be checked as follows:\end{change}.

\newcount\min \min=5
\newcount\max \max=20
%\countUsesVersionControlRepository=4
\ifnum \countUsesVersionControlRepository>\max countUsesVersionControlRepository is far too large! \fi
\ifnum \countUsesVersionControlRepository<\min countUsesVersionControlRepository is far too small! \fi
\ifnum \countUsesVersionControlRepository>\max
	\else 
		\ifnum \countUsesVersionControlRepository<\min
			\else
								\textbf{This is \emph{automatic\/} confirmation that $$\the\min \leq \mbox{\tt countUsesVersionControlRepository} \leq \the\max$$ and therefore falls within pre-defined sanity limits set for this paper.}
		\fi
\fi

@@

\begin{change}
\highlightmysubsection{Defending against system problems}
Code can become obsolete as programming languages develop and compilers are improved. Typically, compilers first warn that code is ``deprecated'' and then later versions reject the old code. Furthermore, when code is run on different computers, different operating systems, and with different compilers, it is common to obtain different results. Data, too, is subject to the same problems, but data standards and formats are far more stable than code standards, so ``data rot'' is less of a risk (but no less a problem when it occurs) than ``software rot.''

Additionally, errors can be the result of human slips, such as accidentally deleting a line of code or a line of data in a spreadsheet. Such corruption errors are hard to detect unless specific steps are taken to ensure the integrity of code and data \cite{excel-fiasco}. Checksums are the simplest way to detect such errors, but during active research more refined techniques might be used in addition, for example checking that the number of rows of data in a spreadsheet monotonically increases. In the present paper, the JSON data is more structured than a spreadsheet matrix, and a number (as it happens, \the\JSONerrorCount) of other consistency checks are imposed on the data.

To protect against version, portability and other problems, the Github repository for the present paper includes a check on software versions and a checksum check for all possibly affected files, including the data file. This does not solve the problem, but it ensures anyone developing or reproducing the paper's work will at least be forewarned of potential version or portability problems. The Github repository itself can be used to restore files that have been corrupted.

\highlightmysubsection{Problems of restrictive journal policies}
\label{restrictive-policies}
Automatically generated variables are used throughout the paper and this \supplement.
As usual, \LaTeX\ detects any spelling errors in the use of variables, thus helping protect the paper against typos that could otherwise mislead the paper's readers. Conveniently, \LaTeX\ also supports sophisticated calculations itself \cite{latex-calculus}, so the typeset paper can use any variable values in further calculations without going back to modify the data source file (in the present case, \texttt{data.js}). In practice this enables the author to avoid copying-and-pasting values from a data source or calculator, and then overlooking keeping them up to date with changes to the data or formula required.

For example, the caption of table \ref{table-repo-summary} in the main paper calculates its ``\plural{\pubdelayinmonths}{month}'' figure from the generated variables recording the repository date of cloning used to provide the data to construct the table. The number of months will of course be correctly updated if the paper's repository \csname cite-PostoperativeOutcomes.RiskNet\endcsname\ is subsequently checked again:

\begin{quote}
At the time of cloning and checking all repositories in \clonewhen, paper \csname cite-PostoperativeOutcomes.RiskNet\endcsname\ still had nothing in its repository except a single file still saying ``...code coming soon...,'' despite \the\pubdelayinmonths\ months having already elapsed since the submitted paper had claimed the code could be accessed in its repository.\end{quote}

Of course, the data generation process itself checks that this surprising statement remains valid, and provides a warning if the wording may need revising.

Unfortunately, although using generated variables and analyses from a paper's data is a very simple technique to help make published papers more reliable, some journals and preprint servers (such as \emph{IEEE Transactions on Software Engineering}, \emph{PLOS ONE}, and \emph{arXiv\/}) do not permit papers to be submitted using \LaTeX\ source code that uses the standard \texttt{\textbackslash input}, \texttt{\textbackslash bibliography}, and other related commands. Typically they also do not support running any data collection or analysis either (which the present paper does when it clones repositories).

Another program (\texttt{expand.js}) was therefore written to recursively expand included files so the expanded version can be submitted adhering to any such restrictive policy. Of course, the expanded version now contains all variables as fixed constants, so the submitted paper is misleading and useless to other researchers if the data is modified --- the effort to ensure all published numbers are automatically correct is defeated. Such restrictive publishing policies undermine reproducibility.
\end{change}

@@

It is noteworthy that none of the journals sampled permit any reliable style of managing data in published papers, such as described above in sections \ref{detecting-against-error} and \ref{restrictive-policies}. In particular, for all the papers that had accessible code, the code included explicit (and relevant) data that was not archived \emph{as\/} data in the journal repositories.
\\

\newdimen\tabwidth
\tabwidth=\textwidth
\advance \tabwidth by -4em
\def\specifyurl{\vskip 3mm
\begingroup\raggedleft\scriptsize
} 
\def\specifyurletc#1{\endgroup\\
\hbox{ }\hfill\begin{change}\begin{tabular}{|p{\tabwidth}@{}}
#1
\end{tabular}\end{change}\vskip .3cm
}

{\sf
\noindent\textbf{Extract from \emph{Royal Society Open Science\/} author guidelines} \\
--- It is a condition of publication that authors make the primary data, materials (such as statistical tools, protocols, software) and code publicly available. These must be provided at the point of submission for our Editors and reviewers for peer-review, and then made publicly available at acceptance. [\ldots] As a minimum, sufficient information and data are required to allow others to replicate all study findings reported in the article. Data and code should be deposited in a form that will allow maximum reuse. As part of our open data policy, we ask that data and code are hosted in a public, recognized repository, with an open licence (CC0 or CC-BY) clearly visible on the landing page of your dataset.}

\specifyurl
\url{royalsociety.org/journals/authors/author-guidelines/#data}
\specifyurletc{Accessed 29 July 2020; the policy has been revised (undated, but accessed 2 February 2022) but retains the same principles; full policy now available via a DOI \cite{RS-policy}. The policy still retains an emphasis on data accessibility, and continues a lack of awareness that code and data are equivalent and often mixed (see section \ref{on-code-data-publication}).}

\noindent\textbf{Extract from \emph{Nature Digital Medicine\/} author guidelines}\\
{\sf --- A condition of publication in a Nature Research journal is that authors are required to make materials, data, code, and associated protocols promptly available to readers without undue qualifications. [\ldots] A condition of publication in a Nature Research journal is that authors are required to make unique materials promptly available to others without undue qualifications.}

\specifyurl
\url{www.nature.com/nature-research/editorial-policies/reporting-standards#availability-of-data}
\specifyurletc{Accessed 29 July 2020; since updated to require [in part] ``Upon publication, Nature Portfolio journals consider it best practice to release custom computer code in a way that allows readers to repeat the published results. Code should be deposited in a DOI-minting repository such as Zenodo, Gigantum or Code Ocean and cited in the reference list following the guidelines described here.'' (accessed 2 February 2022).
}
